---
title: "RWorksheet_Cahutay#2"
author: "Mark Anton B. Cahutay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Number 1: Vector using ":" operator
```{r}
#A. Sequence from -5 to 5
mySequence <- -5:5
mySequence
#The output shows a sequence of numbers starting from negative 5 to positive 5

#B. The value of x is a sequence of numbers from 1 to 7
x <- 1:7
x
```

Number 2: Vector using seq() Function
```{r}
#A. seq(1,3, by=0.2) with specified step size
mySequence2 <- seq(1,3, by = 0.2)
mySequence2
#The output shows a sequence of numbers from 1 iterating by 0.2 until it reaches 3
```

Number 3: Factory Workers
```{r}
#A. Access 3rd element
factory_vector <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 
                    35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 
                    50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 
                    35, 24,33, 41, 53, 40, 18, 44, 38, 41, 48, 
                    27, 39, 19, 30, 61, 54, 58, 26, 18)
factory_vector[3]
#The value of the 3rd element is 22

#B. Access 2nd and 4th element
factory_vector[2]
factory_vector[4]
#The 2nd value holds number 28, while the 4th holds 36

#C. Access all except 4th and 12th
exclude <- factory_vector[-c(4, 12)]
exclude
```

Number 4: Vector names(x)
```{r}
#A. 
x <- c("first" = 3, "second" = 0, "third" = 9)
names(x)

#B.
#The first line of the output of the code below is the strings "first" and "third" and the next line shows the number the strings are holding/equal to.
x[c("first", "third")]

```

Number 5: Sequence x from -3:2
```{r}
x <- -3:2

#A. Modify 2nd element
x[2] <- 0
#The output will generate a sequence of number from -3 to 2, however the second element was changed due to it being reassigned a number = 0.

#B. Code and its output
x
```

Number 6: Diesel Fuel by Mr. Cruz
A. Data frame for month, price per liter, and purchase quantity
```{r}
#R script and its output
diesel_data <- data.frame(
  month = c("Jan", "Feb", "March", "Apr", "May", "June"),
  price = c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00),
  purchase_quantity = c(25, 30, 40, 50, 10, 45)
)
diesel_data

```
Number 6: 
B. The average fuel expenditure of Mr. Cruz from Jan to June
```{r}
#R Script and its output
expenditure <- weighted.mean(diesel_data$price, diesel_data$purchase_quantity)
expenditure
```


Number 7: River Data
A. Type "rivers" in R console
```{r}
rivers
#vector data of rivers
data <- c(length(rivers), 
          sum(rivers),
          mean(rivers), 
          median(rivers), 
          var(rivers), 
          sd(rivers), 
          min(rivers),
          max(rivers)
)
```
Number 7:
B. The results of the rivers data
```{r}
data
```
Number 7:
C. R Script and its output
```{r}
rivers_data <- data.frame(
  Data = c("Length", "Sum", "Mean", "Median", "Varian", "Standard Deviation", "Minimum", "Maximum"),
  Value = data
)
rivers_data
```


Number 8: Celebrity annual pay rankings
A.1 Vectors
```{r}
#Power ranking vector
power_ranking <- c(1:25)
power_ranking
```
A.2
```{r}
#Name vector
celebrity_name <- c(
               "Tom Cruise", "Rolling Stones", "Oprah Winfrey", 
               "U2", "Tiger Woods", "Steven Spielberg", 
               "Howard Stern", "50 Cent", "Cast of the Sopranos", 
               "Dan Brown", "Bruce Springsteen", "Donald Trump", 
               "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John",
               "David Letterman", "Phil Mickelson", "J.K Rowling",
               "Brad Pitt", "Peter Jackson", "Dr. Phil McGraw",
               "Jay Leno", "Celine Dion", "Kobe Bryant"
)
celebrity_name
```
A.3
```{r}
#Pay vector
pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 
         55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31
)
pay
```

Number 8: 
B. Modified J.K. Rowling
```{r}
power_ranking[19] <- 15
pay[19] <- 90

#modified output
power_ranking
celebrity_name
pay
```
Number 8:
C. import excel/csv file
```{r}
forbes_ranking <- read.csv("PowerRanking.csv")
forbes_ranking
```
Number 8: 
D. Access row 10 to 20
```{r}
row_data <- forbes_ranking[10:20, ]

save(row_data, file = "Ranks.RData")

load("Ranks.RData")
row_data

#E. The output generated a tabular like format of excel showing the data from row 10 - 20.
```


Number 9:
A.  Import the excel file to RStudio. 
```{r}
library(openxlsx)
hotel_vienna <- read.xlsx("hotels-vienna.xlsx")
```

B. Dimensions of the data set
```{r}
dim(hotel_vienna)
```  

C. Select Columns
```{r}
columns <- hotel_vienna[, c("country", "neighbourhood", "price", "stars", "accommodation_type", "rating")]
columns
```

D. Saving data as **new.RData 
```{r}
save(columns, file = "new.RData")
```   

E. Display the first six rows and last six rows of the new.RData. What is the R script?
```{r}
load("new.RData")
head(columns)
tail(columns)
```


Number 10: List of 10 vegetables
A.
```{r}
veggies <- list("Eggplpant", "Cabbage", "Broccoli", "Water Spinach", "Squash",
               "Onion", "Pipino", "Patola", "Patani", "Malunggay"
)
veggies
```

B. Addition of 2 veggies
```{r}
veggies <- append(veggies, list("Carrot", "Spinach"))
veggies
```   

C. Added 4 additional veggies
```{r}
veggies <- append(veggies, list("Bitter Gourd", "Tomato", "Potato", "Kasoy"), after = 5)
veggies
length(veggies)
```   

D. Removed vegetables in index 5, 10, and 15.
```{r}
veggies <- veggies[-c(5, 10, 15)]
veggies
length(veggies)
```   
